---
title: "Details"
description: >
  Details.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Details}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  - pdf/R-validation.pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

The qualification process carried out by this package can be split into two steps:

1. **Installation Qualification (IQ)**
2. **Operational Qualification (OQ)**

During IQ, several checks are performed to ensure that the R installation is properly 
set up and configured. During OQ, base and recommended packages are checked to 
ensure they are functioning correctly.

## Installation Qualification (IQ)

The following information is collected for display in the PDF report during the
IQ process:

* R installation location
* R welcome banner
* System and user information
* Platform specific variables
* R software version information
* Numerical characteristics of the machine
* Session information
* Package library location
* Pandoc version information
* version and installation path of TinyTex

## Operational Qualification (OQ)

During OQ, the functions `testInstalledBasic` and `testInstalledPackages` from the
base `tools` package are used to check the functionality of base and recommended packages. 
The following 7 checks are performed:

1. System test execution via `testInstalledBasic("both")`
2. Base package examples execution via `testInstalledPackages(scope="base", types="examples")`
3. Base package vignettes execution via `testInstalledPackages(scope="base", types="vignettes")`
4. Recommended package examples execution via `testInstalledPackages(scope="recommended", types="examples")`
5. Recommended package vignettes execution via `testInstalledPackages(scope="recommended", types="vignettes")`
6. Base package tests execution via `testInstalledPackages(scope="base", types="tests")`
7. Recommended package tests execution via `testInstalledPackages(scope="recommended", types="tests")`

The check result of each package and function are displayed in each section of the PDF report, 
along with any error messages encountered during execution.

In the case of both IQ and OQ, any error will be captured and displayed in the PDF report, 
along with a message indicating that the check has failed.

## Report Example

The following is an example of the PDF report generated by the `rqualify()` function, 
which includes the results of both IQ and OQ checks, as well as a summary table of the checks 
performed and their outcomes.


```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("pdf/R-validation.pdf")
```
