---
geometry: "left=0.5in,right=0.5in,top=1.0in,bottom=1.0in"
header-includes:
  - \let\CheckCommand\providecommand
  - \usepackage{array}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{textcomp}
  - \usepackage{stmaryrd}
  - \usepackage{graphicx}
  - \usepackage{booktabs}
  - \usepackage{fancyhdr}
  - \usepackage[linktocpage=true,colorlinks=true,urlcolor=blue,linkcolor=black]{hyperref}
  - \usepackage[format=hang,font=bf,belowskip=1ex,singlelinecheck=off]{caption}
  - \usepackage[hang,norule,stable,multiple]{footmisc}
  - \usepackage{listings}
  - \lstloadlanguages{R} 
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{R Software IQ OQ}
  - \fancyfoot[C]{\thepage}
  - \fancyfoot[R]{\today}
  - \renewcommand\headrulewidth{0pt}
  - \renewcommand\footrulewidth{0.4pt}
  - \addtocontents{toc}{\protect\thispagestyle{fancy}}
  - \addtocontents{lof}{\protect\thispagestyle{fancy}}
  - \addtocontents{lot}{\protect\thispagestyle{fancy}}
  - \setlength\parindent{0in}
  - \addtolength{\parskip}{\baselineskip} 
urlcolor: blue
output:
  pdf_document: default
---


<!-- LaTeX customizations -->
<!-- 1) Manually define forestgreen -->
<!-- 2) Listings settings for pretty output -->
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}

\lstset{
language=R,
caption=,
basicstyle=\scriptsize\ttfamily,
commentstyle=\ttfamily\color{gray},
numbers=left,
numberstyle=\ttfamily\color{gray}\footnotesize,
stepnumber=1,
numbersep=5pt,
showspaces=false,
showstringspaces=false,
showtabs=false,
xleftmargin=0.25in,
xrightmargin=0.25in,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname,
escapeinside={},
keywordstyle=\ttfamily\color{red},
stringstyle=\ttfamily\color{forestgreen},
morekeywords={OK,NA,PASS,FAIL,FAILED,failed,Execution,halted},
deletekeywords={R,C,for,terms,is,if,file,path,For,Platform,machine,system,status,rev,packages,factor,
package,graphics,methods,class,grid,Version,or,as,double,min,exp,integer,max,var,
q,inherits,report,ifelse,on,help,browser,quit,try,options,plot,ps,save,lines,list,
character,attr,se,~,_,/}
}

\UseRawInputEncoding


```{r codeexec, echo=FALSE}

# Helper function for running code at the command line
code_exec <- function(code_block, file_prefix, folder_output){

  # Set code file name
  file_code <- file.path(folder_output, paste0(file_prefix, ".R"))

  # Write code_block to R file
  fileConn <- file(file_code)
  writeLines(code_block, fileConn)
  close(fileConn)

  # Set output file name
  file_output <- file.path(folder_output, paste0(file_prefix, "Out.txt"))

  # Command to run @ command line
  cmd <- paste(shQuote(file.path(R.home("bin"), "R")),
               "CMD BATCH --vanilla --no-timing", 
               shQuote(file_code), 
               shQuote(file_output))

  # Run the command
  run <- try(system(cmd))

  # Load results and return
  readLines(file.path(folder_output, paste0(file_prefix, "Out.txt")))
}

```



```{r titlepage, echo=FALSE}

# Get R installation facts
RVersionInfo <- R.Version()
Version      <- RVersionInfo$version.string
Arch         <- gsub("_", " ", RVersionInfo$arch)
Platform     <- gsub("_", " ", RVersionInfo$platform)

```

\begin{titlepage}

\begin{center}

\Large Installation and Operational Qualification Testing For R Statistical Software\

\Large `r Version`\

\Large Architecture: `r Arch`\

\Large Platform: `r Platform`\

\vspace{1.5in}

\Large{\emph{
\today
}}

\end{center}

\end{titlepage}


\newpage

\tableofcontents

\newpage

# Introduction

The R software being tested in this report was downloaded from The Comprehensive R Archive Network (CRAN):

<http://cran.r-project.org/>

or a CRAN mirror:

<http://cran.us.r-project.org/mirrors.html>

The R software was installed on this computer in a manner consistent with "The R Installation and Administration Manual" (the Manual) which is available from:

<http://cran.r-project.org/doc/manuals/R-admin.html>

The Manual provides recommendations for post-installation testing of R. The procedures for this testing are described in:

<http://cran.r-project.org/doc/manuals/R-admin.html#Testing-a-Unix_002dalike-Installation>

for Unix, Linux and OSX installations and:

<http://cran.r-project.org/doc/manuals/R-admin.html#Testing-a-Windows-Installation>

for Windows installations. The results contained within this report are based upon an automated implementation of the procedures described in the Manual.

General guidance for the use of R in regulated clinical trials has been provided by the R Foundation in a document entitled:

\begin{center}

R: Regulatory Compliance and Validation Issues

A Guidance Document for the Use of R in Regulated Clinical Trial Environments

\end{center}

which is available from:

<http://www.r-project.org/doc/R-FDA.pdf>

The above document describes various characteristics of R, including the Software Development Life Cycle (SDLC) and relevant aspects of 21 CFR Part 11 compliance issues as they may pertain to the use of R for statistical analysis applications for clinical trials.

The output on the following pages of this report describe various technical characteristics of the computer upon which R is running, the R installation, current R session information and is followed by a series of tests for the so-called "Base" and "Recommended" packages which are a part of the official R distribution, as released by the R Foundation. Any installed packages that are not "Base" or "Recommended" packages are not part of this validation procedure and are ignored.



